---
title: "Group Project 1"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Ian Mawn, Kaitlin McLain, and Emily Scholfield
date: 21 Feb 2022
---

Welcome to this R tutorial for Bio208! This is designed for a very beginner to R, so we are going to start with the very basics.

Typically, when doing data analysis in R, we want to begin by loading in the data and packages that we will use.

1. Loading data and libraries

R has many built in functions that will allow us to do perform all kinds of actions. However, sometimes it is useful to load some additional functions that other R users have created. We do this by loading in libraries. In the coade chunk below (the grey box, this is where code is typed and run) you can see that I have loaded in some libraries.

```{r Load Libraries, include=FALSE}
# Load other packages here.
#I just put in any libraries we might need. Don't know if we actually need them, or need more.
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("readxl")) install.packages("readxl"); library(readxl)
```
Libraries:
For some examples, one library is called tidyverse. It contains a ton of useful tools, but the one most relevant to this tutorial is called ggplot. ggplot allows us to create all kinds of plots to visualize our data, which you will learn more about below. You can also see the library called cowplot. cowplot is useful as it allows us to easily change the aesthetics of our graphs, and keep the designs consistent between different graphs. Finally, you can see the library readxl. This is useful for exactly what it sounds like... reading Microsoft Excel files! This bring us to the next step, entering data into R.

Data:
In the code chunk below I am going to read in some data regarding features of different species of iris (the flower). There are a few main parts of the code to think about when loading in data. First, the read_csv() function tells R to load in the comma separated values (csv) file, and the name of the file goes in the parentheses. Here, you can see that our data is called "iris.csv". It is very important to type the file name exactly as it is saved on your computer, otherwise R will not be able to find it. Once the data is loaded in, we need to call it something. That is what the "IrisData <-" part of the code means. In English, it translates to: IrisData is the data found in the csv file "iris.csv".

```{r echo=FALSE}
IrisData <- read_csv ("iris.csv")
IrisData
```

2. Using R

Now that we have the pieces we need, we can start learning about R. 




3. Exploring data





4. Graphs/Visualization





5. Hypothesis Testing with Statistical Tests





6. Drawing Conclusions




























This project will require you to develop a tutorial to teach Bucknell students how to use R for graphing and data analysis. 

## Target Audience

Discuss with your group the target audience for the tutorial. 
Examples could be one of the new core Biology classes, another 300-level course (not 364), or a research group. 

```{r echo=FALSE}
IrisData <- read_csv ("iris.csv")
```

## Groups

You will work with the same groups as for the Homework 2 Peer review. 

```{r}
PeerGroups1 <- read_csv ("PeerGroups1.csv")
PeerGroups1 %>% 
  select(-MAJR, -CLAS, ) %>%
  print()
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Identify and obtain suitable dataset
- Use a Github repository and version control to collaborate on the project
- Spend 4-6 hours preparing, coding, and testing tutorial
  + Data exploration
  + Data visualization
  + Hypothesis testing
- Present tutorial in class
- Provide public archive suitable for sharing to students/faculty

Tutorials from previous classes can be viewed at our public github site: https://github.com/Bucknell-Biol364

Each group should use an *Acknowledgements* section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for the development of a tutorial that goes beyond the minimal expectations listed above.

## Sample Dataset

One of the possible datasets to use for the tutorial can be found in the datasauRus package.

```{r}
datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarize(
      mean_x    = mean(x),
      mean_y    = mean(y),
      std_dev_x = sd(x),
      std_dev_y = sd(y),
      corr_x_y  = cor(x, y)
      )
```

Boxplots of either the x or the y value show that there is something strange going on.

```{r}
ggplot(datasaurus_dozen, aes(x = x, colour = dataset))+
    geom_boxplot()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol = 3)
```

```{r}
ggplot(datasaurus_dozen, aes(x = y, colour = dataset))+
    geom_boxplot()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol = 3)
```

But you have to visualize all of the data with a scatter plot to really see the patterns.

```{r fig.height=12, fig.width=9}
  ggplot(datasaurus_dozen, aes(x = x, y = y, colour = dataset))+
    geom_point()+
    theme_void()+
    theme(legend.position = "none")+
    facet_wrap(~dataset, ncol = 3)
```

And did you notice the code in the {r} codechunk header that controlled the size of the output in the Rmd? Pretty neat trick!

And here are two versions of the data that you could use in your data visualization tutorial. 
To use them you would probably want to change the names of the datasets and also make x and y more meaningful. 
Then save them as a csv or tsv to be imported later for your tutorial. 

```{r}
datasaurus_long <- datasaurus_dozen
datasaurus_wide <- datasaurus_dozen_wide
head(datasaurus_long)
head(datasaurus_wide)
```

# Acknowledgements

DatasauRus package and description below: Stephanie Locke https://github.com/jumpingrivers/datasauRus

The datasauRus package wraps the awesome Datasaurus Dozen dataset, which contains 13 sets of x-y data. Each sub-dataset has five statistics that are (almost) the same in each case. (These are the mean of x, mean of y, standard deviation of x, standard deviation of y, and Pearson correlation between x and y). However, scatter plots reveal that each sub-dataset looks very different. The dataset is intended to be used to teach students that it is important to plot their own datasets, rather than relying only on statistics.

The Datasaurus was created by Alberto Cairo in this great [blog post](http://www.thefunctionalart.com/2016/08/download-datasaurus-never-trust-summary.html).

He's been subsequently made even more famous in the paper [Same Stats, Different Graphs: Generating Datasets with Varied Appearance and Identical Statistics through Simulated Annealing](https://www.autodeskresearch.com/publications/samestats) by Justin Matejka and George Fitzmaurice.

In the paper, Justin and George simulate a variety of datasets that the same summary statistics to the Datasaurus but have very different distributions. 

This package also makes these datasets available for use as an advanced [Anscombe's Quartet](https://en.wikipedia.org/wiki/Anscombe%27s_quartet), available in R as `anscombe`.
